---
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
editor_options: 
  chunk_output_type: console
---
```{r set-options, echo=FALSE, cache=FALSE}
options(width = 2000)
```
# Discussion 8

Page 312

**Excercise #3 Write a program to toss a coin 10,000 times. Let Sn be the number of heads in the first n tosses. Have your program print out, after every 1000 tosses, Sn − n/2. On the basis of this simulation, is it correct to say that you can expect heads about half of the time when you toss a coin a large number of times?**

I chose to use both `sample` and `rbinom` as functions for generating a random number that simulates a coin flip. The $S_{n}$ value is calculated using `sample` and the $S_{x}$ is using `rbinom`

```{r, warning=FALSE,message=FALSE}
library(tidyverse)
```
```{r}
set.seed(1234)
coin <- c('heads','tails')
n <- 1
sn <- 0
sx <- 0
df <- data.frame(matrix(ncol = 5, nrow = 0))
colnames(df) <- c('n', 'error', 'sn','binomError','sx')
repeat {
  # get a value of 'heads' or 'tails' using sample
  result <- sample(coin,1, replace = TRUE, prob = c(.5,.5)) 
  # get a value of 0 or 1 using rbinom
  r <- rbinom(1,1,prob = .5)
  
  if(result == 'heads'){
    sn <- sn + 1
  }
  if(r == 1){
    sx <- sx + 1
  }
  error <- sn - (n/2)
  binomError <- sx - (n /2)
  
  df<-df %>% add_row(n, error, sn,binomError,sx)
  
  if(n %% 1000 == 0){
     print(paste('sn:',error, ' sx:', binomError))
  }
  
  n <- n + 1
  #break after 10,000 iterations
  if(n > 10000){
    break
  }
}

ggplot(df, aes(n)) + 
  geom_line(aes(y = error, colour = "sample error")) + 
  geom_line(aes(y = binomError, colour = "binom error")) +
  ylab('Sn - n/2')

df <- df %>% mutate(rSn = sn / n) %>% mutate(rSx = sx / n)

df %>% ggplot(aes(n)) + 
  geom_line(aes(y=rSn, colour='ratio w/ sample')) + 
  geom_line(aes(y=rSx, colour='ratio w/ binom')) + ylab('heads to tails ratio')
```


**Yes, though what I believe to be the "chance" factor calculated by the $S_{n} - \frac{n}{2}$ appears to oscillate wildly as demonstrated by my plot, the plot of the ratio of heads to tails shows that there is a rapid convergence towards 0.5 using both methods of generating a random number.**