---
title: "DATA 605 - Discussion 14"
author: "Nick Oliver"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
editor_options: 
  chunk_output_type: console
---

# Discussion 14 Section - 8.8 Exerecise 26 (pg 496)

**In Exercises 25 – 30, use the Taylor series given in Key Idea 8.8.1 to create the Taylor series of the given functions.**

26. $f(x) = e^{-x}$

From 8.8.1 the Taylor series approximation of $e^x$ is

$$
e^x = \sum_{n=0}^{\infty} \frac{x^n}{n!} = 1 + x + \frac{x^2}{2!} + \frac{x^3}{3!} + ...
$$
To create the Taylor series for $f(x) = e^{-x}$  replace $x$ with $-x$
$$
e^{-x} = \sum_{n=0}^{\infty} \frac{-x^n}{n!} = 1 - x + \frac{x^2}{2!} - \frac{x^3}{3!} + ...
$$
I wrote a function to generate the Taylor series
```{r}
taylorSeries <- function(f){
  function(x){
    res <- 1
    n <- 1
    while(n <= f){
      c <- (((-x)^n) / factorial(n))
      res <- res + c
      n <- n + 1
    }
    res
  }
}
```
Plotting $e^{-x}$ in red then each approximation in darker shades of blue to visualize the approximation getting better
```{r}
curve(expr = exp(1)^-x, -2,2, col='red')
for(i in 1:5){
  t <- taylorSeries(i)
  curve(t, -2,2, col=rgb(0,0,(51 * i)/255,.4),add=T)
}

```


```{r}

```