---
title: "DATA 605 - Discussion 11"
author: "Nick Oliver"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
editor_options: 
  chunk_output_type: console
---

# Discussion 11

I am currently volunteer teaching Python to high school students and the tool we are using makes the usage data available as a CSV. Wanted to do a linear regression on if the amount of time students spend in the tool correlates with their quiz scores.

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(kableExtra)
library(performance)
```
## Loading Data
```{r}
df <- read.csv('https://raw.githubusercontent.com/nolivercuny/DATA605/main/week%2011/data/student_data.csv')
```
## Cleanup & Aggregation

Filter out instructors, sum the minutes in the tool and get the mean of their quiz scores
```{r}
df <- df %>%
  filter(!grepl("instructor",username))
df <- df %>%
  mutate(studyMinsSum = rowSums(select(df,2:226),na.rm = T)) %>%
  select(c(username,
           studyMinsSum,starts_with("Unit.1.Quiz") |
             starts_with("Unit.2.Quiz") |
             starts_with("Unit.3.Quiz")))
df <- df %>%
  mutate(quizMean = rowMeans(select(df,3:5),na.rm=T))
```

## Analysis

Plotting students times vs scores and showing the raw data
```{r}
plot(df$quizMean ~ df$studyMinsSum, main="Effort vs. Scores",
xlab="Total Work Time (m)", ylab="Mean Quiz Score")
df %>% 
  kable()
```
Use built in `lm` function to fit linear model. Summary shows a `p-value` of $\approx$ 0.68 which much greater than .05 which indicates the correlation is not statistically significant. 
```{r, warning=FALSE}
model <- lm(data=df, quizMean ~ studyMinsSum)
summary(model)
```
Using ggplot and the `stat_smooth` there does appear to be a very slight positive correlation
```{r, warning=FALSE}
ggplot(data = df, aes(x = studyMinsSum, y = quizMean)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE)
```

Residuals plot

```{r}
 res = resid(model)
plot(df$studyMinsSum, res,
     ylab="Residuals", xlab="Study Time")
 abline(0, 0)                  # the horizon
ggplot(data = df, aes(x = model$residuals)) +
    geom_histogram(fill = 'steelblue', color = 'black', binwidth = 1) +
    labs(title = 'Histogram of Residuals', x = 'Residuals', y = 'Frequency')
```

Shamelessly stealing the idea of using the `performance` library's `check_model` function as well.

```{r}
check_model(model)
```

## Summary
Does it meet the conditions of linear regression?

**Linearity: No. Based on the point plot there really does not appear to be a linear relationship.**

**Nearly normal residuals: Yes. We can observe from the histogram plot of the residuals that their distribution sort of resembles the normal distribution**

**Constant variability: Yes. We can observe from the scatter plot of the residuals that there is no pattern present which indicates that there is constant variability**

**Independent observations: Yes. The study time and scores between individuals is independent**

Is the linear model appropriate?

**Hard to say. Doesn't seem to meet all the criteria of linear regression as I understand it. This is data from a single class of 17 students with only 3 quizzes taken so far. I think there probably just isn't enough data to draw any significant conclusions**

```{r}
cdf = read.csv('https://online.stat.psu.edu/stat462/sites/onlinecourses.science.psu.edu.stat462/files/data/poverty/index.txt', sep = '\t' )

cor(cdf[,!(names(cdf) %in% c('Brth15to17','Brth18to19','TeenBrth','Location'))], cdf$TeenBrth)
```
