---
title: "DATA605 - Assignment 11"
author: "Nick Oliver"
output:
  pdf_document:
    latex_engine: xelatex
    extra_dependencies: ["geometry", "multicol", "multirow"]
editor_options: 
  chunk_output_type: console
---

# Assignment 11

## Question

Using the “cars” dataset in R, build a linear model for stopping distance as a function of speed and replicate the analysis of your textbook chapter 3 (visualization, quality evaluation of the model, and residual analysis.)

## Solution

Loaded the `tidyverse` library
```{r, echo=FALSE, message=FALSE}
library(tidyverse)
```
View the data set to find the column names
```{r}
glimpse(cars)
```
Use the `lm` function to fit a linear regression model
```{r}
model <- lm(data = cars, dist ~ speed)
```

### Visualizing the data

As the text does you can use the buil-in plot function to plot the scatter plot then add the regression line from generated by the `lm` function using `abline`
```{r}
plot(cars)
abline(model)
```

You can also use `ggplot` to plot the scatter plot of speed vs stopping distance. Use `stat_smooth` to add a linear fitted line to the plot representing a linear regression model.
```{r}
cars %>% ggplot(aes(x = speed, y = dist)) +
  geom_point() +
  stat_smooth(method = "lm", se = T) +
  xlab('Speed') +
  ylab('Stopping Distance')
```

### Evaluating the quality of the model

```{r}
summary(model)
```

Standard error at least five times to ten times smaller than corresponding coefficients?
$3.9324 / 0.4155 \approx 9.464$ Yes it is about 9 times smaller