---
title: "DATA605 - Assignment 15"
author: "Nick Oliver"
output:
  pdf_document:
    latex_engine: xelatex
    extra_dependencies: ["geometry", "multicol", "multirow"]
editor_options: 
  chunk_output_type: console
---
```{r, echo=FALSE, message=FALSE}
library(tidyverse)
```
# Assignment 15

## 1.

Find the equation of the regression line for the given points. Round any final values to the nearest hundredth, if necessary. (5.6,8.8), (6.3,12.4), (7,14.8), (7.7,18.2), (8.4,20.8)

**Answer:** $y = -14.8 + 4.26 \times x$


```{r}
x<-c(5.6, 6.3, 7, 7.7, 8.4)
y<-c(8.8, 12.4,14.8,18.2,20.8)
df <- data.frame(x,y)
df %>% ggplot(aes(x=x,y)) +
  geom_point() +
stat_smooth(method = "lm", se = FALSE)
summary(lm(y~x))
```



## 2.

Find all local maxima, local minima, and saddle points for the function given below. Write your answer(s) in the form ( x, y, z ). Separate multiple points with a comma.
$f(x,y)=24x-6xy^2 -8y^3$

Find derivatives with respect to x and y

```{r}
f <- expression(24*x - 6 * x * y^2 - 8*y^3)
D(f,'x')
D(f,'y')
```

$\frac{\partial}{\partial{x}} = 24 - 6y^2$
$\frac{\partial}{\partial{y}} =-12xy - 24y^2$

Set derivatives equal to 0 to find critical points

$ 0 = 24 - 6y^2 \rightarrow y = \pm 2$

Solve for x when $y=2$

$0 =-12x*2 - 24*2^2 \rightarrow x = -4$

Solve for x when $y=-2$

$0 =(-12x*-2) - 24*(-2)^2 \rightarrow x = 4$

Obtain z values by plugging in values for x,y

```{r}
x<--4
y<-2
print(paste0('z=',24*x - 6 * x * y^2 - 8*y^3))

x<-4
y<--2
print(paste0('z=',24*x - 6 * x * y^2 - 8*y^3))
```

***Critical points () and ()***

## 3.

A grocery store sells two brands of a product, the "house" brand and a "name" brand. The manager estimates that if she sells the "house" brand for x dollars and the "name" brand for y dollars, she will be able to sell $81 - 21x + 17y$ units of the "house" brand and $40 + 11x - 23y$ units of the "name" brand.

Step 1. Find the revenue function $R(x,y)$.

Step 2. What is the revenue if she sells the "house" brand for $2.30 and the "name" brand for $4.10?

## 4. 

A company has a plant in Los Angeles and a plant in Denver. The firm is committed to produce a total of 96 units of a
product each week. The total weekly cost is given by $C(x, y) = \frac{1}{6}x^2 + \frac{1}{6}y^2 + 7x + 25y + 700$, where x is the number of
units produced in Los Angeles and y is the number of units produced in Denver. How many units should be produced in each plant to minimize the total weekly cost?

## 5. 

Evaluate the double integral on the given region.
   $$\iint_R (e^{8x+3y})dA;R:2\le x \le 4 and 2 \le y \le 4$$
  
Write your answer in exact form without decimals.