---
title: "DATA605 - Assignment 4"
author: "Nick Oliver"
output:
  pdf_document:
    latex_engine: xelatex
    extra_dependencies: ["geometry", "multicol", "multirow"]
editor_options: 
  chunk_output_type: console
---

# Homework 4

## Problem

With the attached data file, build and visualize  eigenimagery that accounts for 80% of the variability.  Provide full R code and discussion.

## Solution

I borrowed heavily from Diego Herrero's Face Recognition using Eigenfaces[^1] and also used code from the Wikipedia page on Eigenfaces[^2]

## Load Libraries

Only using OpenImageR library to load images
```{r}
library(OpenImageR)
```

## Load images and extract data into a matrix

```{r}
basePath <- '~/jpg/'
files <- list.files('~/jpg')
data <- matrix(0, length(files), prod(c(1200,2500,3))) 

for (i in 1:length(files)) {
  img <- readImage(paste0(basePath,files[i])) 
  r  <- as.vector(img[,,1])
  g  <- as.vector(img[,,2])
  b  <- as.vector(img[,,3])
  
  data[i,] <- t(c(r, g, b))
}
```

## Convert vectors into a dataframe
```{r}
shoes <- data.frame(x = t(data))
```
## scale and center
```{r}
scaled <- scale(shoes, center = TRUE, scale = TRUE)
```
## Calculate covariance
```{r}
sig <- cov(scaled)
```

## Compute eigenvectors and values
```{r}
eig <- eigen(sig)
eigenvalues  <- eig$values
eigenvectors <- eig$vectors
```

## Evaluate the number of principal components needed to represent 80% Total variance.

```{r}
eigsum <- sum(eigenvalues);
csum <- 0
k80 <- 0
tv <- 0
for(i in 1:17){
   csum <- csum + eigenvalues[i]
    tv <- csum / 17
    print(tv)
    if(tv > 0.80){
      k80 <- i
      break
    }
}
k80
```

```{r}
shoesScaled <- diag(eigenvalues[1:k80]^(-1/2)) / (sqrt(nrow(scaled)-1))
eigenShoes <- scaled %*% eigenvectors[,1:k80] %*% shoesScaled
```

```{r}
  eigenimage <- array(eigenShoes[,3], c(1200,2500,3))  
  imageShow(eigenimage)
```


## References

1. https://rpubs.com/dherrero12/543854/

2. https://en.wikipedia.org/wiki/Eigenface

[^1]:https://rpubs.com/dherrero12/543854/

[^2]:https://en.wikipedia.org/wiki/Eigenface