---
title: "DATA 605 - Discussion 12"
author: "Nick Oliver"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
editor_options: 
  chunk_output_type: console
---

# Discussion 12

I am using the state of Delaware's open data set. The regression model is predicting the percentage of students who are proficient on the MATH portion of the "Smarter Balanced Summative Assessment" in 2021

PctProficient: The percent of students who earned an achievement level considered proficient. This is the number of students who earned a proficient achievement level divided by the number of students who tested.

Soruce: https://data.delaware.gov/Education/Student-Assessment-Performance/ms6b-mt82

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(kableExtra)
library(performance)
library(MASS)
```
## Loading Data
```{r}
df <- read.csv('https://raw.githubusercontent.com/nolivercuny/DATA605/main/week%2012/Filtered_Student_Assessment_Performance.csv')

```

## Cleanup & Aggregation

This is a very large dataset (over 700k rows) I pre-filtered it so it is easier to work with. Removed all redacted data, removed data for the entire state, removed data without race and gender information and kept just data for the MATH assessment. 

Dropping rows that add no extra information

```{r}
fdf <- df %>%
  mutate(NonWhite = Race != 'White') %>% 
  mutate(Male = Gender == 'Male') %>%
  mutate(GradeNum = as.numeric(gsub("([0-9]+).*$", "\\1", Grade))) %>%
  mutate(NonFosterCare = SpecialDemo == 'Non-Foster Care') %>%
  mutate(NonLowIncome = SpecialDemo == 'Non Low-Income') %>%
  mutate(LowIncome = SpecialDemo == 'Low-Income') %>%
  mutate(ActiveEL = SpecialDemo == 'Active EL Students') %>%
  mutate(NonHomeless = SpecialDemo == 'Non-Homeless') %>%
  mutate(NonELStudents = SpecialDemo == 'Non-EL Students') %>%
  mutate(Disabilities = SpecialDemo == 'Students with Disabilities') %>%
  mutate(MilitaryConnected = SpecialDemo == 'Military Connected Youth') %>%
  dplyr::select(NonWhite,Male,GradeNum,NonFosterCare,NonLowIncome,LowIncome,ActiveEL,NonHomeless,NonELStudents,Disabilities,MilitaryConnected,Tested,PctProficient)
 
glimpse(fdf)
```

## Analysis

Use built in `lm` function to fit linear model. Summary shows a `p-value` of $\approx$ 2e-16 which much less than .05 which indicates the correlation is statistically significant. Also found out that `~ .` is shorthand for including all columns as predictors.
```{r, warning=FALSE}
model <- lm(data=fdf, PctProficient ~ .)
summary(model)
```

I experimented with the `stepAIC` function from the `MASS` library which can do both backward and forward predictor elimination using the Akaike Information Criterion (AIC) to compare the fit of the model. Interestingly it appears to drop the NonLowIncome and MilitaryConnected predictors. 
```{r}
stepModel <- stepAIC(model, direction = 'both')
summary(stepModel)
```

$$PctProficient = 55.202  -13.63248 \times NonWhiteTRUE \\ 
+ 2.11625 \times MaleTRUE \\
- 1.78763  \times GradeNum \\
-2.35069 \times NonFosterCareTRUE \\
-12.83572 \times    LowIncomeTRUE \\
-7.78963 \times   ActiveELTRUE \\
-2.23635 \times NonHomelessTRUE  \\
-3.33112 \times NonELStudentsTRUE \\
-20.23805  \times DisabilitiesTRUE\\
-0.09309  \times Tested$$

What this means according to the data is you start at a baseline proficiency of 55.202% subtract for non-white, add for male, subtract for grade number, subtract if non-foster care, subtract if they are low income, subtract if they are active el, subtract if they are non-homeless, subtract if they are non-el, subtract if they are disabilities, and subtract based on the number of students tested.

Residuals plot

```{r}
 res = resid(stepModel)
plot(df$PctProficient, res,
     ylab="Residuals", xlab="% Proficient")
 abline(0, 0)                  # the horizon
ggplot(data = df, aes(x = stepModel$residuals)) +
    geom_histogram(fill = 'steelblue', color = 'black', binwidth = 1) +
    labs(title = 'Histogram of Residuals', x = 'Residuals', y = 'Frequency')
```

Using the `performance` library's `check_model` function as well.

```{r}
check_model(stepModel)
```

## Summary

Is the linear model appropriate?

**The linear model appears to be appropriate. The check model function and plot of the residuals indicates this is an appropriate model**

