---
title: "DATA605 - Assignment 12"
author: "Nick Oliver"
output:
  pdf_document:
    latex_engine: xelatex
    extra_dependencies: ["geometry", "multicol", "multirow"]
editor_options: 
  chunk_output_type: console
---
```{r, echo=FALSE, message=FALSE}
library(tidyverse)
```
# Assignment 12

The attached who.csv dataset contains real-world data from 2008. The variables included follow.
Country: name of the country
LifeExp: average life expectancy for the country in years
InfantSurvival: proportion of those surviving to one year or more
Under5Survival: proportion of those surviving to five years or more
TBFree: proportion of the population without TB.
PropMD: proportion of the population who are MDs
PropRN: proportion of the population who are RNs
PersExp: mean personal expenditures on healthcare in US dollars at average exchange rate
GovtExp: mean government expenditures per capita on healthcare, US dollars at average exchange rate TotExp: sum of personal and government expenditures.

## 1. 

**Provide a scatterplot of LifeExp~TotExp, and run simple linear regression. Do not transform the variables. Provide and interpret the F statistics, R^2, standard error,and p-values only. Discuss whether the assumptions of simple linear regression met.**

```{r}
df <- read.csv('https://raw.githubusercontent.com/nolivercuny/DATA605/main/week%2012/who.csv')
plot(df$LifeExp ~ df$TotExp,
xlab="Total Expenditures", ylab="Life Expectancy")
model <- lm(data=df, LifeExp ~ TotExp)
summary(model)
```

**F-statistic:**65.26 is meaningless for single parameter models

$R^2$:0.2577 indicates a relatively poorly fitting model as it is closer to 0 than 1

**Standard Error**:7.795e-06 Standard error  6.297e-05 /7.795e-06 = 8.078255. The large ratio indicates little variability in the slope estimate

**P-value**:7.714e-14 much smaller than 0.05 which indicates statistical significance

### Are the assumptions of a simple linear regression met?

(1) linearity:No model is not linear

(2) nearly normal residuals: Yes residuals look nearly normal based on histogram plot


```{r}
ggplot(data = df, aes(x = model$residuals)) +
    geom_histogram(fill = 'steelblue', color = 'black') +
    labs(title = 'Histogram of Residuals', x = 'Residuals', y = 'Frequency')
```

(3) constant variability: No, an obvious pattern in the fitted vs residual plot

```{r}
ggplot(data = model, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("Fitted values") +
  ylab("Residuals")
```


## 2. 

**Raise life expectancy to the 4.6 power (i.e., LifeExp^4.6). Raise total expenditures to the 0.06 power (nearly a log transform, TotExp^.06). Plot LifeExp^4.6 as a function of TotExp^.06, and r re-run the simple regression model using the transformed variables. Provide and interpret the F statistics, R^2, standard error, and p-values. Which model is "better?"**

```{r}
df <- df %>%
  mutate(LifeExpFourSix = LifeExp^4.6) %>%
  mutate(TotalExpPointSix = TotExp^.06)
model <- lm(data=df, LifeExpFourSix ~ TotalExpPointSix)
summary(model)
```

**F-statistic:**507.7 is meaningless for single parameter models

$R^2$:0.7298 indicates a well fitting model as it is much closer to 1 than 0

**Standard Error**:27518940 Standard error  620060216 /27518940 = 22.53213. The large ratio indicates little variability in the slope estimate

**P-value**:2.2e-16 much smaller than 0.05 which indicates statistical significance

The summary statistics all indicate that the second model is a better fitting model.

## 3.

**Using the results from 3, forecast life expectancy when TotExp^.06 =1.5. Then forecast life expectancy when TotExp^.06=2.5.**

$LifeExpectancy = -736527910  + 620060216 \times TotExp$

```{r}
l1 <- -736527910  + (620060216 * 1.5)
l1
l1 ^ (1/4.6)
l2 <- -736527910  + 620060216 * 2.5
l2
l2 ^ (1/4.6)
```

Answer: $\approx$ 63 year life expectancy with 1.5 total expenditures and $\approx$ 87 life expectancy with 2.5 total expendeitures. 

## 4. 

**Build the following multiple regression model and interpret the F Statistics, R^2, standard error, and p-values. How good is the model?**
LifeExp = b0+b1 x PropMd + b2 x TotExp +b3 x PropMD x TotExp

```{r}
multipleModel <- lm(data=df,LifeExp ~ PropMD + TotExp + (PropMD * TotExp))
summary(multipleModel)
```
**F-statistic:**34.49

$R^2$:0.3574 indicates a well fitting model as it is much closer to 1 than 0

**Standard Error**:

2.788e+02  1.497e+03 / 2.788e+02 = 5.36944

8.982e-06 7.233e-05 / 8.982e-06 = 8.052772

1.472e-03  -6.026e-03/ 1.472e-03  =-4.09375

All of the standard error ratios are near 5-10 times smaller than the coefficients

**P-value**:2.2e-16 much smaller than 0.05 which indicates statistical significance

## 5. 

**Forecast LifeExp when PropMD=.03 and TotExp = 14. Does this forecast seem realistic? Why or why not?**

$LifeExpectancy = 6.277e+01  + 1.497e+03 \times PropMD + 7.233e-05 \times TotExp - -6.026e-03 \times PropMD \times TotExp$

```{r}
PropMD <- 0.03
TotExp <- 14
6.277e+01  + (1.497e+03 * PropMD) + (7.233e-05 * TotExp) -(6.026e-03 * PropMD * TotExp)
```

Life expectancy of 107.6785 years Seems a little unrealistic as I would assume it is relatively uncommon to live to the age of 107. 