---
output:
  pdf_document:
    latex_engine: xelatex
    extra_dependencies: ["geometry", "multicol", "multirow"]
editor_options: 
  chunk_output_type: console
---

# Discussion 7

Page 203

## 38 A manufactured lot of buggy whips has 20 items, of which 5 are defective. A random sample of 5 items is chosen to be inspected. Find the probability that the sample contains exactly one defective item

### (a) if the sampling is done with replacement.

**If the sampling is done with replacement this can be solved using binomial**

$$P(X = k) = \binom{n}{k}p^k(1-p)^{n-k}$$
```{r}
p <- 5/20
n <- 5
k <- 1
nChooseK <- factorial(n) / (factorial(k) * factorial(n -k))

(nChooseK * p ^ k * (1 - p)^(n -k))
#check answer
dbinom(1,5,.25)
```

### (b) if the sampling is done without replacement.

**Without replacement then we use a hypergeometric formula**

$$P(X = k) = \frac{\binom{K}{k}\binom{K-k}{n-k}}{\binom{N}{n}}$$
N is the population size, K is the number of success states in the population, n is the number of draws (i.e. quantity drawn in each trial), k is the number of observed successes.

```{r}
K <- 5
k <- 1
N <- 20
n <- 5
(choose(K,k) * choose(N - K,n - k)) / choose(N,n)
#check answer
dhyper(1,5,15,5)
```
