---
title: "DATA605 - Assignment 7"
author: "Nick Oliver"
output:
  pdf_document:
    latex_engine: xelatex
    extra_dependencies: ["geometry", "multicol", "multirow"]
editor_options: 
  chunk_output_type: console
---

# Assignment 7 

## 1. Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi's. Find the distribution of Y .



## 2. Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has a manufacturer's expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R Code for each part.).

### a. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a geometric. (Hint: the probability is equivalent to not failing during the first 8 years..)

### Probability
$$Pr(X = x) = (1-p)^{k}p$$
```{r}
geomProb <- (1 - (.1))^(8) * .1
geomProb
dgeom(8,.1)
```

Answer: **0.04304672**

### Expected Value

$$ E[X] = \frac{1}{p} $$
```{r}
1/geomProb
```
$$Std Dev = \frac{\sqrt{1-p}}{p}$$
```{r}
sqrt(1 - geomProb) / geomProb
```

### b. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as an exponential.

### Probability

$$Pr(X=k) = \lambda e^{-\lambda x }$$
```{r}
prob <- .1
prob * exp(-prob * 8)
dexp(8,.1)
```
Answer: **0.0449329**

### Expected Value

$$E[x] = \frac{1}{\lambda}$$
```{r}
1 / .1
```

Answer: **10**

### Standard Deviation

$$StdDev = \sqrt{\frac{1}{\lambda^2}}$$
```{r}
sqrt(1/(.1^2))
```

Answer: **10**

### c. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8 years)

### Probability

$$Pr(X = k) = \binom{n}{k}p^k(1-p)^{n-k}$$
```{r}
k <- 0
prob <- 1/10
n <- 8
nCk <- choose(n,k)
 (nCk * (prob ^ k)) * (1 - prob) ^ (n-k)
dbinom(k,n,prob)
```

Answer: **0.4304672**

### Expected Value

$$E[x] = np$$

Answer: **0.8**

### Standard Deviation

$$StdDev = \sqrt{np(1-p)}$$
```{r}
prob <- 1/10
n <- 8
sqrt(n * prob * ( 1 - prob))
```
Answer: **0.8485281**

### d. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a Poisson

### Probability

$$Pr(X=k) = \frac{\lambda^ke^{-\lambda}}{k!}$$
```{r}
k <- 0
prob <- 1/10

lambda <- 8 * prob / 1

((lambda ^ k) * exp(-lambda))/factorial(k)
dpois(0,.8)
```

Answer: **0.449329**

### Expected Value

$$E[X] = \lambda $$

Answer: **0.8**

### Standard Deviation

$$StdDev = \sqrt{\lambda}$$
```{r}
sqrt(lambda)
```
Answer: **0.8944272**

