---
title: "DATA605 - Assignment 9"
author: "Nick Oliver"
output:
  pdf_document:
    latex_engine: xelatex
    extra_dependencies: ["geometry", "multicol", "multirow"]
editor_options: 
  chunk_output_type: console
---

# Assignment 9

## 1. 11 page 363
11 The price of one share of stock in the Pilsdorff Beer Company (see Exercise 8.2.12) is given by $Y_n$ on the nth day of the year. Finn observes that the differences $X_n = Y_{n+1} − Y_n$ appear to be independent random variables with a common distribution having mean $\mu = 0$ and variance $\sigma^2 = 1/4$. If $Y_1 = 100$, estimate the probability that $Y_{365}$ is

```{r}
n <- 364
y1 <- 100
variance  <- 1/4
mu <- 0
sd <- sqrt(n * variance)
```

**(a) $\ge$ 100.**

$Pr(Y_{365} \ge 100) = 0.5$

```{r}
pnorm(100 - y1, mu, sd, lower.tail = F)
```

**(b) $\ge$ 110.**

$Pr(Y_{365} \ge 110) = 0.1472537$

```{r}
pnorm(110 - y1, mu, sd, lower.tail = F)
```

**(c) $\ge$ 120.**

$Pr(Y_{365} \ge 120) = 0.01801584$

```{r}
pnorm(120 - y1, mu, sd, lower.tail = F)
```

## 2. Calculate the expected value and variance of the binomial distribution using the moment generating function.

From Wikipedia the $M_{x}(t)$ for a binomial distribution is
	$${\displaystyle \left(1-p+pe^{t}\right)^{n}}$$
The mean is equal to the first derivative of the moment generating function at t=0 and the variance is equal to the second derivative of the moment generating function at t=0 minus the mean squared

**Mean**
$$
\begin{split}
\mu = M'(0) = n(1-p + pe^0)^{n-1}(pe)^0 \\
= n(1-p+p)^{n-1}p(1) \\
= np
\end{split}
$$

**Variance**

$$
\begin{split}
M''(0) = n(1-p+pe^0)^{n-1}(pe^0)+(pe^0)n(n-1)(1-p+pe^0)^{n-2}pe^0 \\
= n(1)^{n-1}p+pn(n-1)(1)^{n-2}p(1) \\
= np + n^2p^2 - np^2
\end{split}
$$
$$
\begin{split}
\sigma^2 = M''(0) - M'(0)^2 = np + n^2p^2-np^2 - np^2 \\
= np + n^2p^2 - np^2 - n^2p^2 \\
= np - np^2
\end{split}
$$

## 3. Calculate the expected value and variance of the exponential distribution using the moment generating function.

From Wikipedia the $M_{x}(t)$ for a exponential distribution is

$$
{\displaystyle \left(1-t\lambda ^{-1}\right)^{-1},~t<\lambda } \\
=\frac{\lambda}{\lambda-t}
$$

The mean is equal to the first derivative of the moment generating function at t=0 and the variance is equal to the second derivative of the moment generating function at t=0 minus the mean squared

**Mean**
$$
\begin{split}
\mu = M'(0) = \frac{\lambda}{(\lambda-0)^2} \\
= \frac{\lambda}{\lambda^2} \\
= \frac{1}{\lambda}
\end{split}
$$

**Variance**

$$
\begin{split}
M''(0) = \frac{2\lambda}{(\lambda-0)^3} \\
= \frac{2}{\lambda^2}
\end{split}
$$
$$
\begin{split}
\sigma^2 = M''(0) - M'(0)^2 = \frac{2}{\lambda^2} - (\frac{1}{\lambda})^2 \\
= \frac{1}{\lambda^2}
\end{split}
$$