---
title: "DATA605 - Assignment 3"
author: "Nick Oliver"
output:
  pdf_document:
    latex_engine: xelatex
    extra_dependencies: ["geometry", "multicol", "multirow"]
editor_options: 
  chunk_output_type: console
---


IS 605 FUNDAMENTALS OF COMPUTATIONAL MATHEMATICS - 2014


1. Problem set 1 

(1) What is the rank of the matrix A?

\[
A = \begin{bmatrix}
1&2&3&4\\
−1&0&1&3\\
0&1&−2&1\\
5&4&−2&−3
\end{bmatrix}
\]

Convert A to RREF

\[
A = \begin{bmatrix}
1&2&3&4\\
−1&0&1&3\\
0&1&−2&1\\
5&4&−2&−3
\end{bmatrix} \xrightarrow{RREF} 
\begin{bmatrix}
1&0&0&0\\
0&1&0&0\\
0&0&1&0\\
0&0&0&1
\end{bmatrix}
\]

Rank is 4. There are no linear independent rows

```{r}
A = matrix(c(1,2,3,4,-1,0,1,3,0,1,-2,1,5,4,-2,-3),nrow=4,byrow = TRUE)
A
pracma::rref(A)
Matrix::rankMatrix(A)
```

(2) Given an mxn matrix where m > n, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero?

The maximum rank would be `m` as it is possible as in problem set one that no rows are linearly independent. The minimum rank is 1

(3) What is the rank of matrix B?
\[
B = \begin{bmatrix}
1&2&1\\
3&6&3\\
2&4&2
\end{bmatrix}
\]

Convert B to reduced row echelon form. Count number of non-zero rows

\[
B = \begin{bmatrix}
1&2&1\\
3&6&3\\
2&4&2
\end{bmatrix} \xrightarrow{RREF}
\begin{bmatrix}
1&2&1\\
0&0&0\\
0&0&0
\end{bmatrix}
\]
```{r}
B = matrix(c(1,2,1,3,6,3,2,4,2),nrow=3,byrow=TRUE)
B
pracma::rref(B)
Matrix::rankMatrix(B)
```

2. Problem set 2
Compute the eigenvalues and eigenvectors of the matrix A. You’ll need to show your work. You’ll need to write out the characteristic polynomial and show your solution.

\[
A = \begin{bmatrix}
1&2&3\\
0&4&5\\
0&0&6
\end{bmatrix}
\]

\[
\lambda{I_{3}} = \begin{bmatrix}\lambda&0&0\\0&\lambda&0\\0&0&\lambda\end{bmatrix}
\]

\[
\lambda{I_{3}} - A = 
\begin{bmatrix}
\lambda-1 & -2 & -3 \\
0 & \lambda-4 & -5 \\
0 & 0 & \lambda-6
\end{bmatrix}
\]

Apply rule of Sarrus

\[
\lambda{I_{3}} - A = 
\begin{bmatrix}
\lambda-1 & -2 & -3 \\
0 & \lambda-4 & -5 \\
0 & 0 & \lambda-6
\end{bmatrix}
\begin{matrix}
\lambda-1 & -2 \\
0 & \lambda-4 \\
0 & 0
\end{matrix}
\]

\[
(\lambda-1 )(\lambda-4)(\lambda-6) + (-2)(-5)(0) + (-3)(0)(0) - (-2)(0)(\lambda-6) - (\lambda-1)(-5)(0) - (-3)(\lambda-4)(0)
\]

(3) Please show your work using an R-markdown document. Please name your assignment submission with your first initial and last name.
