---
title: "DATA 605 - Discussion 13"
author: "Nick Oliver"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
editor_options: 
  chunk_output_type: console
---

# Discussion 13

Using R, provide the solution for any exercise in either Chapter 4 or Chapter 7 of the calculus textbook. 
If you are unsure of your solution, post your concerns.

14. A powerline is to be run to an offshore facility in the manner described in Example 4.3.3. The offshore facility is 5 miles at sea and 2 miles along the shoreline from the power plant. It costs \$50,000 per mile to lay a power line underground and $80,000 to run the line underwater.

How much of the power line should be run underground to minimize the overall costs?

Cost = 50,000 * land + 80,000 * water

We choose $x$ to represent the distance on water, meaning the distance on land is 2 miles minus x

$Cost(x) = 50000 \times (2 - x)  + 80000 \times \sqrt{x^2 + 5^2}$

Find the minimum $Cost(2)$

```{r}
50000 * (2 - 2)  + 80000 * sqrt(2^2 + 5^2)
```

Find the maximum $Cost(0)$
```{r}
50000 * (2 - 0)  + 80000 * sqrt(0^2 + 5^2)
```

Find $Cost'(x)$
```{r}
cx <- expression(50000 * (2 - x)  + 80000 * sqrt(x^2 + 25))
d <- D(cx, "x")
d
```
$\frac{80000x}{\sqrt{25 + x^2}} -50000 = 0$

Solving for $x \approx 4$

Plug 4 in for c(x)

```{r}
(2 - 4)

```
