---
title: "DATA 605 - Discussion 13"
author: "Nick Oliver"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
editor_options: 
  chunk_output_type: console
---

# Discussion 13

13. A power line is to be run to an offshore facility in the manner described in Example 4.3.3. The offshore facility is 2 miles at sea and 5 miles along the shoreline from the power plant. It costs \$50,000 per mile to lay a power line underground and $80,000 to run the line underwater.

How much of the power line should be run underground to minimize the overall costs?

Cost = 50,000 * land + 80,000 * water

We choose $x$ to represent the distance on water, meaning the distance on land is 5 miles minus x

$Cost(x) = 50000 \times (5 - x)  + 80000 \times \sqrt{x^2 + 2^2}$

Find $Cost'(x)$
```{r}
cx <- expression(50000 * (5 - x)  + 80000 * sqrt(x^2 + 4))
d <- D(cx, "x")
d
```
$\frac{80000x}{\sqrt{x^2 + 4}} -50000 = 0$

Solving for x yields $x \approx 1.6$

Find the minimum underground distance by simply subtracting 5 by 1.6
```{r}
(5 - 1.6)
```
**Approximately 3.4 miles should be underground.**