---
title: "DATA605 - Assignment 13"
author: "Nick Oliver"
output:
  pdf_document:
    latex_engine: xelatex
    extra_dependencies: ["geometry", "multicol", "multirow"]
editor_options: 
  chunk_output_type: console
---
```{r, echo=FALSE, message=FALSE}
library(tidyverse)
```
# Assignment 13

## 1
Use integration by substitution to solve the integral below.

### Answer

$$
\begin{split}
\int{4e^{-7x}dx} \\
u = -7x \\
du = -7dx \Rightarrow dx = \frac{du}{-7}
\end{split}
$$
by substitution
$$
\begin{split}
\int{4e^u \frac{du}{-7}} \\
-\frac{4}{7}\int{e^udu}\\
-\frac{4}{7}e^u + C\\
-\frac{4}{7}e^{-7x} + C\\
\end{split}
$$


## 2

Biologists are treating a pond contaminated with bacteria. The level of contamination is changing at a rate of
$\frac{dN}{dt} = -\frac{3150}{t^4}-220$ bacteria per cubic centimeter per day, where t is the number of days since treatment began. Find a function N( t ) to estimate the level of contamination if the level after 1 day was 6530 bacteria per cubic centimeter.

### Answer

We know the rate of change is the derivative of the function $N(t)$ so to find $N(t)$ we take the integral
$$
\begin{split}
\frac{dN}{dt} = N'(t) = \frac{3150}{t^4}-220 \\
N(t) = \int{\frac{3150}{t^4}-220dt} \\
 = \frac{1050}{t^3}-220t + C
\end{split}
$$
We are given $N(1) = 6530$ so we can plug that in to solve for $C$

$$
\begin{split}
6530 = \frac{1050}{1^3}-220 * 1 + C \\
6530 = 830 + C \\
C = 5700
\end{split}
$$
This gives us

$$
N(t) = \frac{1050}{t^3}-220t + 5700
$$
## 3

Find the total area of the red rectangles in the figure below, where the equation of the line is f ( x ) = 2x - 9.

### Answer

By looking at the graph the limits are observed as 4.5 as the lower bound and 8.5 as the upper bound. Using R to perform the integration
```{r}
f <- function(x){2*x - 9}
integrate(f, 4.5, 8.5)
```

**16 is the area of the rectangle**

## 4

Find the area of the region bounded by the graphs of the given equations.
$y=x^2-2x-2, y=x+2$

### Answer

First plot the curves to find the intersection

```{r}
funOne <- function(x){x^2 - 2*x - 2}
funTwo <- function(x){x+2}
curve(funOne, -2,5)
curve(funTwo,-2,5,add=T)
```

From the plot of the curves the intersection occurs at x=-1 and x=4 which are our lower and upper bounds. To find the area we take the integral of the upper function minus the lower function

$$
\begin{split}
\int_{-1}^{4}{(x^2 - 2x -2) - (x+2)dx} \\
\int_{-1}^{4}{x^2 - 3x -4dx}
\end{split}
$$

```{r}
funDiff <- function(x){x^2 - 3*x - 4}
integrate(funDiff, -1,4)
```

Area is $\approx$ -20.833

## 5

A beauty supply store expects to sell 110 flat irons during the next year. It costs $3.75 to store one flat iron for one year. There is a fixed cost of $8.25 for each order. Find the lot size and the number of orders per year that will minimize inventory costs.

### Answer

To minimize inventory costs we need to create a function $C(x)$ using what was given then take its derivative. 

- Costs 3.75 to store one iron per year
- Each order costs 8.25
- Expect to sell 110 per next year

$C(x) = 8.25x + \frac{3.75}{2x}$

```{r}
D(expr = expression(8.25 * (110/x) + 3.75* (x/2)), 'x')
```

$C'(x) = 3.75 * (1/2) - 8.25 * (110/x^2)$

Setting $C'(x) = 0$ you get $x = 22$, which means by plugging 22 in for the number of unknown orders 110/22 = 5 orders will minimize costs

## 6
Use integration by parts to solve the integral below.
$$
\int{ln(9x)x^6dx}
$$

### Answer

Integration by parts 
$$
\begin{split}
\int{udv} = uv - \int{vdu} \\
u = ln(9x),du = \frac{1}{x},v = \frac{1}{7}x^7,dv = x^6\\
ln(9x) \frac{1}{7}x^7 - \int{\frac{1}{7}x^7\frac{1}{x}dx} \\
ln(9x) \frac{1}{7}x^7 - \frac{1}{7}\int{x^6dx} \\
ln(9x) \frac{1}{7}x^7 - \frac{1}{7} \times \frac{1}{7}x^7 + C \\
ln(9x) \frac{1}{7}x^7 - \frac{x^7}{49} + C
\end{split}
$$

## 7
Determine whether f ( x ) is a probability density function on the interval $\brack{1,e^6}$. If not, determine the value of the definite integral $f(x) = \frac{1}{6x}$

### Answer

In order for f(x) to be a PDF the integral must be = 1

```{r}
fX <- function(x){1/(6*x)}
integrate(fX, 1,exp(1)^6)
```
