---
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
editor_options: 
  chunk_output_type: console
---

# Discussion 10

Page 422 11.2 Problem #6

**6 In the Land of Oz example (Example 11.1), change the transition matrix by making R an absorbing state. This gives**

$$
P = \begin{array}{cc} &
\begin{array}{ccc} R & N & S\end{array}
\\
\begin{array}{ccc}
R \\
N \\
S\end{array}
&
\left(
\begin{array}{ccc}
1 & 1 & 0 \\
1/2 & 0 & 1/2 \\
1/4 & 1/4 & 1/2\end{array}
\right)\end{array}
$$
**Find the fundamental matrix N, and also Nc and NR. Interpret the results.**


1) Convert P to canonical form

$$
\begin{split}
P ^n = \left(\begin{array}{c|c} 
	Q & R\\ 
	\hline 
	0 & I 
\end{array}\right) 
\\
=  \begin{array}{cc} &
\begin{array}{ccc} N & S & R\end{array}
\\
\begin{array}{ccc}
N \\
S \\
R\end{array}
&
\left(
\begin{array}{cc|c}
0 & 1/2 & 1/2 \\
1/4 & 1/2 & 1/4 \\
\hline
0 & 0 & 1\end{array}
\right)\end{array}
\end{split}
\\

$$

$$
\begin{split}
Q = \left(\begin{array}{cc} 
	0 & 1/2\\ 
	1/4 & 1/2
\end{array}\right) 

\\
R = \left(\begin{array}{c} 
	1/2\\ 
	1/4
\end{array}\right) 
\end{split}
$$

3) According to the book $N = (I - Q)^{-1}$

```{r}
Q <- matrix(c(0,1/2,1/4,1/2), ncol = 2, byrow=TRUE)

I <- diag(2)
N <- solve(I - Q) 
N
```

4) Computing Nc. According to the book c is a column vector whose entries are all equal to 1
```{r}
Nc <- N %*% matrix(c(1,1))
Nc
```

5) Computing NR, using R from above
```{r}
R <- matrix(c( 0.5, 0.25 ), ncol = 1, byrow = TRUE)

NR <- N %*% R
NR
```

```{r}
library(expm)
P <- matrix(c(.50, .25, .25, .50, 0, .50, .25, .25, .50), ncol=3,nrow=3, byrow = TRUE) 
n<- 10
u1 <- c(0,1,0)

compute_un <- function(n,u,P){
  Pn <- P %^% n 
  u %*% Pn
}

compute_un(10,u1,P)
```