---
title: "DATA605 - Assignment 10"
author: "Nick Oliver"
output:
  pdf_document:
    latex_engine: xelatex
    extra_dependencies: ["geometry", "multicol", "multirow"]
editor_options: 
  chunk_output_type: console
---

# Assignment 10

Smith is in jail and has 1 dollars; he can get out on bail if he has 8 dollars. A guard agrees to make a series of bets with him. If Smith bets A dollars, he wins A dollars with probability .4 and loses A dollars with probability .6. Find the probability that he wins 8 dollars before losing all of his money if

**(a) he bets 1 dollar each time (timid strategy).**

A fit's the form of the Gambler's Ruin problem from the text "A gambler starts with a “stake” of size s. She plays until her capital reaches the value M or the value 0."

s = 1
M = 8

$$
q_z = \frac{\frac{q}{p}^z - 1} {\frac{q}{p}^M - 1}
$$

```{r}
z <- 1
M <- 8
q <- .6
p <- .4
qz <- ((q/p)^z -1) / ((q/p)^M -1)
qz
```

Answer: There is a 0.0203 probability that Smith wins the game without running out of money

**(b) he bets, each time, as much as possible but not more than necessary to bring his fortune up to 8 dollars (bold strategy).**

Using the bold strategy there are now only transition states

Starting with one dollar:

bets 1 dollar then .4 wins 1 dollar $\rightarrow$ 2 dollars || .6 loses 1 dollar $\rightarrow$ 0 dollars

bets 2 dollars then .4 wins 2 dollars $\rightarrow$ 4 dollars || .6 loses 2 dollars $\rightarrow$ 0 dollars

bets 4 dollars then .4 wins 4 dollars $\rightarrow$ 8 dollars || .6 loses 4 dollars $\rightarrow$ 0 dollars
  
You can model this with an absorbing state markov chain

$$
P = 
\left(\begin{array}{ccccc} 
	1 & 0 & 0 & 0 & 0 \\ 
	0.6 & 0 & 0.4 & 0 & 0 \\
	0.6 & 0 & 0 & 0.4 & 0 \\ 
	0.6 & 0 & 0 & 0 & 0.4 \\
	0 & 0 & 0 & 0 & 1
\end{array}\right)
$$

You can use Theorem 11.2 from the book $u^{(n)} = uP^n$ where n = 3
```{r, echo=FALSE, message=FALSE}
library(expm)
```
```{r}
#using library expm
P <- matrix(c(
  c(1,0,0,0,0),
  c(0.6,0,0.4,0,0),
  c(0.6,0,0,0.4,0),
  c(0.6,0,0,0,0.4),
  c(0,0,0,0,1)),nrow = 5, byrow=TRUE)
P
n <- 3
u1 <- c(0,1,0,0,0)
Pn <- P %^% n 
u1 %*% Pn
```

Answer: There is a 0.064 probability that Smith ends up with 8 dollars before losing all his money

**(c) Which strategy gives Smith the better chance of getting out of jail?**

The bold strategy has a higher probability 
