---
title: "DATA605 - Disc 5"
author: "Nick Oliver"
output:
  pdf_document:
    latex_engine: xelatex
    extra_dependencies: ["geometry", "multicol", "multirow"]
editor_options: 
  chunk_output_type: console
---

# 7 A die is rolled until the first time T that a six turns up. 

## (a) What is the probability distribution for T ?

From Wikipedia[^1] "The geometric distribution gives the probability that the first occurrence of success requires k independent trials, each with success probability p. If the probability of success on each trial is p, then the probability that the kth trial (out of k trials) is the first success is $Pr(X=k)=(1-p)^{k-1}p$"

Or as the book defines it $P(T = n) = q^{n−1}p$ where $q = 1 - p$



Probability of rolling a six on a fair die is $Pr(6) = \frac{1}6$

Plot using the density function
```{r}
plot(dgeom(1:25,1/6))
```


$$
\\
$$

$$\\
$$




## (b) Find P (T > 3)

$$P(T > 3) = (1 - \frac{1}{6})^{3 - 1}\frac{1}{6}$$
```{r}
((1 - (1/6)) ^ (4 -1)) * (1/6)
dgeom(3, 1/6)
```




## (c) Find P (T > 6|T > 3)

From the book $P(T >r+s|T > s) = q^s$ where $q=1-p$
```{r}
(1-(1/6))^3
```

[^1]:https://en.wikipedia.org/wiki/Geometric_distribution/
