---
title: "DATA605 - Assignment 5"
author: "Nick Oliver"
output:
  pdf_document:
    latex_engine: xelatex
    extra_dependencies: ["geometry", "multicol", "multirow"]
editor_options: 
  chunk_output_type: console
---

# Assignment 5

## 1. (Bayesian).

A new test for multinucleoside-resistant (MNR) human immunodeficiency virus type 1
(HIV-1) variants was recently developed. The test maintains 96% sensitivity, meaning that, for those
with the disease, it will correctly report “positive” for 96% of them. The test is also 98% specific,
meaning that, for those without the disease, 98% will be correctly reported as “negative.” MNR HIV-1
is considered to be rare (albeit emerging), with about a .1% or .001 prevalence rate. Given the
prevalence rate, sensitivity, and specificity estimates, what is the probability that an individual who is
reported as positive by the new test actually has the disease? If the median cost (consider this the best
point estimate) is about $100,000 per positive case total and the test itself costs \$1000 per
administration, what is the total first-year cost for treating 100,000 individuals?

### what is the probability that an individual who is reported as positive by the new test actually has the disease?


Found an equation for computing positive predictive value using sensitivity, prevalence, specificity 

PPV = (Sensitivity * Prevalence)/[(Sensitivity * Prevalence) + ((1 - Specificity) * (1 - Prevalence))][^1]

```{r}
Sensitivity <- .96
Prevalence <- .001
Specificity <- .98
ppv <- (Sensitivity * Prevalence)/((Sensitivity * Prevalence) + ((1 - Specificity) * (1 - Prevalence)))
ppv * 100
```

### If the median cost (consider this the best point estimate) is about $100,000 per positive case total and the test itself costs \$1000 per administration, what is the total first-year cost for treating 100,000 individuals?

The prevalence of .1% means in 100,000 individuals 100 would be expected to have the disease. Our test has a sensitivity of 96% so 96/100 of those individuals would be expected 

## 2. (Binomial).

The probability of your organization receiving a Joint Commission inspection in any given
month is .05. What is the probability that, after 24 months, you received exactly 2 inspections? What is
the probability that, after 24 months, you received 2 or more inspections? What is the probability that
your received fewer than 2 inspections? What is the expected number of inspections you should have
received? What is the standard deviation?

$P(k) = \binom{n}{k} p^k (1-p)^{n-k}$


###  What is the probability that, after 24 months, you received exactly 2 inspections?
```{r}
dbinom(2, size=24, prob=0.05) 
```
### What is the probability that, after 24 months, you received 2 or more inspections?

```{r}
dbinom(4, size=48, prob=0.05) 
```

### What is the probability that your received fewer than 2 inspections?

$P(x < 2) = P(x = 0) + P (x = 1)$

```{r}
pZero = dbinom(0, size=24, prob = 0.05)
pOne = dbinom(1, size=24, prob = 0.05)
print(paste(pZero,pOne, sep=' + '))
pZero + pOne
```

## 3. (Poisson). 

You are modeling the family practice clinic and notice that patients arrive at a rate of 10 per
hour. What is the probability that exactly 3 arrive in one hour? What is the probability that more than
10 arrive in one hour? How many would you expect to arrive in 8 hours? What is the standard
deviation of the appropriate probability distribution? If there are three family practice providers that can
see 24 templated patients each day, what is the percent utilization and what are your recommendations?

## 4. (Hypergeometric). 

Your subordinate with 30 supervisors was recently accused of favoring nurses. 15
of the subordinate’s workers are nurses and 15 are other than nurses. As evidence of malfeasance, the
accuser stated that there were 6 company-paid trips to Disney World for which everyone was eligible.
The supervisor sent 5 nurses and 1 non-nurse. If your subordinate acted innocently, what was the
probability he/she would have selected five nurses for the trips? How many nurses would we have
expected your subordinate to send? How many non-nurses would we have expected your subordinate to
send?

## 5. (Geometric).

The probability of being seriously injured in a car crash in an unspecified location is about
.1% per hour. A driver is required to traverse this area for 1200 hours in the course of a year. What is
the probability that the driver will be seriously injured during the course of the year? In the course of 15
months? What is the expected number of hours that a driver will drive before being seriously injured?
Given that a driver has driven 1200 hours, what is the probability that he or she will be injured in the
next 100 hours?

## 6. 

You are working in a hospital that is running off of a primary generator which fails about once in 1000
hours. What is the probability that the generator will fail more than twice in 1000 hours? What is the
expected value?

## 7. 

A surgical patient arrives for surgery precisely at a given time. Based on previous analysis (or a lack of
knowledge assumption), you know that the waiting time is uniformly distributed from 0 to 30 minutes.
What is the probability that this patient will wait more than 10 minutes? If the patient has already
waited 10 minutes, what is the probability that he/she will wait at least another 5 minutes prior to being
seen? What is the expected waiting time?

## 8. 

Your hospital owns an old MRI, which has a manufacturer’s lifetime of about 10 years (expected value).
Based on previous studies, we know that the failure of most MRIs obeys an exponential distribution.
What is the expected failure time? What is the standard deviation? What is the probability that your
MRI will fail after 8 years? Now assume that you have owned the machine for 8 years. Given that you
already owned the machine 8 years, what is the probability that it will fail in the next two years? 

[^1]:https://www.ncbi.nlm.nih.gov/books/NBK430867/#:~:text=PPV%20%3D%20(sensitivity%20x%20prevalence),x%20(1%20%E2%80%93%20prevalence))%20%5D
