---
title: "DATA605 - Assignment 2"
author: "Nick Oliver"
output:
  pdf_document:
    latex_engine: xelatex
    extra_dependencies: ["geometry", "multicol", "multirow"]
editor_options: 
  chunk_output_type: console
---

# Assignment 2

## Problem set 1

(1) Show that \[A^TA \neq AA^T\] in general. (Proof and demonstration.)
(2) For a special type of square matrix A, we get AT A = AAT . Under what conditions
could this be true? (Hint: The Identity matrix I is an example of such a matrix).
Please typeset your response using LaTeX mode in RStudio. 

### (1)

\[
A =
\begin{bmatrix}
0&1\\0&1
\end{bmatrix}
\\
A^T = 
\begin{bmatrix}
0&0\\1&1
\end{bmatrix}
\\
A * A^T =  \begin{bmatrix}
1&1\\1&1
\end{bmatrix}
\\
A^T * A =  \begin{bmatrix}
0&0\\0&2
\end{bmatrix}
\\
A * A^T \neq A^T * A
\]
```{r}
A = matrix(c(0,1,0,1), byrow = TRUE, nrow = 2)
A
tA = t(A)
tA

A %*% tA
tA %*% A
```

### (2)

A "diagonal" matrix follows the property \[A*A^T = A^T * A\] A diagonal matrix is a square matrix with 0s outside of the main diagonal. The identity matrix is a trivial example of a diagonal matrix.





\[
A = \begin{bmatrix}6&0&0\\0&2&0\\0&0&4\end{bmatrix}
\\
A^T = \begin{bmatrix}6&0&0\\0&2&0\\0&0&4\end{bmatrix}
\\
A * A ^ T = \begin{bmatrix}36&0&0\\0&4&0\\0&0&16\end{bmatrix}
\\
A^T * A = \begin{bmatrix}36&0&0\\0&4&0\\0&0&16\end{bmatrix}
\\
A * A^T = A^T * A
\]



## Problem set 2

Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being factorized. Radars that track flights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your flight using radars.

Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer. 
You don’t have to worry about permuting rows of A and you can assume that A is less than 5x5, if you need to hard-code any variables in your code. If you doing the entire assignment in R, then please submit only one markdown document for both the problems.

### Function Definition
```{r}
#Doolittle Algorithm
factorizeIntoLU <- function(A) {
  size <- dim(A)[1] #assuming square matrix
  
  L <- diag(0,size)
  U <- diag(0,size)

  for (i in seq(1,size)) {
    for (k in seq(i, size)) {
			acc <- 0;
			for (j in seq(1,i)) {
				acc <- acc + (L[i, j] * U[j, k]);
			}
			U[i, k] <- A[i, k] - acc;
		}
    
		for (k in seq(i,size)){
			if (i == k){
				L[i, i] <- 1
			}
			else {
				acc <- 0;
				for (j in seq(1,i)){
					acc <- acc + (L[k, j] * U[j, i]);
				}
				L[k, i] <- (A[k, i] - acc) / U[i, i];
			}
		}
  }
print('L')
print(L)
print('U')
print(U)
}
```

#### Test Case 1
```{r}

A <- matrix(c(1,2,4,3,8,14,2,6,13),nrow = 3, byrow=TRUE)
A
factorizeIntoLU(A)
```
#### Comparison With Library (1)
```{r}
matrixcalc::lu.decomposition(A)
```
#### Test Case 2
```{r}
A <- matrix(c(1,4,-3,-2,8,5,3,4,7),nrow = 3, byrow=TRUE)
A
factorizeIntoLU(A)
```
#### Comparison With Library (2)
```{r}
matrixcalc::lu.decomposition(A)

```